### 2017-11-15 Latex/meeting1Nov.tex
Adding figures, literature and latex folders

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[a4paper,top=1.5cm,bottom=1.5cm,left=1.5cm,right=1.5cm,%
bindingoffset=0mm]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\newtheorem{prop}{Proposition}
\newtheorem{lemma}{Lemma}
\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.75em depth0em\fi}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{float}
\linespread{1.3}
\raggedbottom

%
\font\reali=msbm10 at 12pt
% subsets of real numbers
\newcommand{\real}{\hbox{\reali R}}
\newcommand{\realp}{\hbox{\reali R}_{\scriptscriptstyle +}}
\newcommand{\realpp}{\hbox{\reali R}_{\scriptscriptstyle ++}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator{\E}{\mathbb{E}}
%

\begin{document}

\Large{

\section*{Results}

\begin{itemize}
	\item Implemented Barsky\&Sims' strategy $\Rightarrow$ Results are in line with the paper. In this case, we can just recover the news shock so we cannot say anything about IT productivity shocks.
	\item Implemented Ryan's strategy $\Rightarrow$ The news shock is in line with B\&S (despite imposing the zero long-run restriction on relative IT prices). However the IT shock explain a minuscule amount of TFP variation and IRFs are largely insignificant.
	\item Implemented a variant of Ryan's strategy ("one step" maximization)
	$$
	\max_{\gamma_{news},\gamma_{IT}} FEV_{TFP,news} + FEV_{TFP,IT}
	$$ s.t. same long-run restriction on news. $\Rightarrow$ Now all the explanation power goes to IT shocks and none is left to news. 
\end{itemize}

\section*{Problems}

\begin{itemize}
	\item The comparison between the two Ryan's strategy variants shows that the data attribute all the variation to either i) the shock that is less restricted or ii) the shock that comes first in the maximization. Our takeaway is that the long-run constraint is not really affecting the result of B\&S. There are two possible explanations: 1) no room for IT shocks. 2) The long-run constraint is not able to disentangle IT from news.
\end{itemize}

\section*{Where to go}

\begin{itemize}
\item An additional constraint?

\item Other thoughts?
\end{itemize}

}
\end{document}