### 2017-11-28 Latex/MacroLunch_Dec2017/macrolunch_dec2017.tex
deleted some old figures and started doing macro lunch presentation

\documentclass{beamer}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{verbatim}

%
\font\reali=msbm10 at 12pt
% subsets of real numbers
\newcommand{\numberset}{\mathbb}
\newcommand{\real}{\hbox{\reali R}}
\newcommand{\N}{\numberset{N}}
\newcommand{\realp}{\hbox{\reali R}_{\scriptscriptstyle +}}
\newcommand{\realpp}{\hbox{\reali R}_{\scriptscriptstyle ++}}
\newcommand{\virgolette}[1]{``#1''}
%

\author[Brianti, Gati]{\\Marco Brianti and Laura Gati}

\institute[Boston College]{Boston College}

\title{Title}

\date{December 13, 2017}

\usetheme{Warsaw}

\begin{document}

\begin{frame}

\maketitle

\end{frame}

%% Slide 1 / Comments I
\begin{frame}
	\frametitle{A primitive question}

	\

	\

	Why has TFP slowed down?

\end{frame}

\end{document}