### 2018-01-09 Latex/Meeting_Ryan_10Jan_2018/meeting10Jan2018.tex
some work on counterfactual

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[a4paper,top=1.5cm,bottom=1.5cm,left=1.5cm,right=1.5cm,%
bindingoffset=0mm]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\newtheorem{prop}{Proposition}
\newtheorem{lemma}{Lemma}
\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.75em depth0em\fi}
\usepackage{tikz}
\usepackage{graphicx}
%\graphicspath{{./Figures/}}
\def \ourFigPath {../../} % this should move up two steps 
% . = current directory
 % .. = parent directory
\usepackage{rotating}
\usepackage{float}
\linespread{1.3}
\raggedbottom

%
\font\reali=msbm10 at 12pt
% subsets of real numbers
\newcommand{\real}{\hbox{\reali R}}
\newcommand{\realp}{\hbox{\reali R}_{\scriptscriptstyle +}}
\newcommand{\realpp}{\hbox{\reali R}_{\scriptscriptstyle ++}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator{\E}{\mathbb{E}}
%

\title{Counterfactuals from SVARs - Questions}
\author{Marco Brianti\\Laura Gati}
\date{10 Jan 2018}

\begin{document}

\maketitle

\

\section{What we have done so far:}
\begin{enumerate}
\item Recovered structural shocks from our identified VAR
\item Used them together with the impulse responses to generate what Kilian calls ``cumulative contribution of a shock $j$" to TFP. These are historical decompositions of TFP in the sense that, relative to the trend of our TFP series, we can see the percent deviations caused by the shocks - together with the responses to them - in our sample. Mathematically, the cumulative effect of a shock $j$ on TFP is:

\begin{equation}
\hat{y}^{j}_{TFP, t} = \sum^{t-1}_{i=0} B(L)^i A s_{j, t-i}
\end{equation}

where $j = \{IT, news\}$ and $s$ is the structural shock. Then, the overall cumulative effect of all shocks is

\begin{equation}
\hat{y}_{TFP, t} = \sum_{j= IT,news} \hat{y}^{j}_{TFP, t}
\end{equation}

\item The object we're interested in is
\begin{equation}
TFP - f\big( \hat{y}^{IT}_{TFP, t} \big)
\end{equation}

and our question is: what is $f$?

\

In other words: the problem comes from the fact that $\hat{y}^{IT}_{TFP, t}$ is in percentage deviations from the trend of TFP. What we're interested in, however, is how TFP would have evolved (in levels) absent the shocks to IT starting in 2000-Q3.

\end{enumerate}

\end{document}