### 2018-02-14 Latex/Drafts/draft_13feb2018.tex
steady state in mathematica, a bibliography and the frame of a paper-style draft for Peter Ireland

\documentclass[11pt]{article}
\usepackage{amsmath, amsthm, amssymb,lscape, natbib}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{multirow}
\usepackage{setspace}
\usepackage[final]{pdfpages}
\usepackage[left=2.5cm,top=2.5cm,right=2.5cm, bottom=2.5cm]{geometry}
\usepackage{natbib} 
\usepackage{bibentry} 
\newcommand{\bibverse}[1]{\begin{verse} \bibentry{#1} \end{verse}}
\newcommand{\vs}{\vspace{.3in}}
\renewcommand{\ni}{\noindent}
\usepackage{xr-hyper}
\usepackage[]{hyperref}

\definecolor{citec}{rgb}{0,0,.5}
\definecolor{linkc}{rgb}{0,0,.6}
\definecolor{bcolor}{rgb}{1,1,1}
\hypersetup{
%hidelinks = true
  colorlinks = true,
  urlcolor=linkc,
  linkcolor=linkc,
  citecolor = citec,
  filecolor = linkc,
  pdfauthor={Laura G\'ati},
}

% A first draft in a paper-format

\geometry{left=.97in,right=1in,top=1in,
bottom=1in}
\linespread{1.5}
\renewcommand{\[}{\begin{equation}}
\renewcommand{\]}{\end{equation}}

\begin{document}

\linespread{1.0}

\title{Title}
\author{Marco Brianti, Laura G\'ati\thanks{%
Department of Economics, Boston College, Chestnut Hill, MA 02467, U.S.A. Email: gati@bc.edu} \\
%EndAName
 Boston College \\
%EndAName
}
\date{Preliminary and Incomplete \\ \today}
\maketitle

%%%%%%%%%%%%%%%%%%%%             ABSTRACT           %%%%%%%%%%%%%%%%%% 
%\begin{abstract}

%\end{abstract}

\newpage
%%%%%%%%%%%%%%%%%%%%               INTRODUCTION                 %%%%%%%%%%%%%%%%%% 
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%               MODEL                 %%%%%%%%%%%%%%%%%% 
\section{Model}

%%%%%%%%%%%%%%%%%%%%               EMPIRICS                 %%%%%%%%%%%%%%%%%% 
\section{Empirics}
\subsection{Data and specification}
\subsection{Results}

%%%%%%%%%%%%%%%%%%%%               CONCLUSION                 %%%%%%%%%%%%%%%%%% 
\section{Conclusion}

 %%%%%%%%%%%%%%%%%%           BIBLIOGRAPHY            %%%%%%%%%%%%%%%%%% 
\newpage
\nocite{*}
\bibliography{itgpt} % try one from BibDesk -- it works!!! 
\bibliographystyle{cell} % cell is the best so far, acm 2nd best

\end{document}