### 2018-04-26 Latex/Slides_Presenter_April2018/dw23jan2018.tex
Preparing slides for the macro reading presenter

\documentclass{beamer}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{float}
\usepackage{csquotes}
\usepackage{sansmathaccent}
\usepackage{subfigure}
\usepackage{multicol}
\pdfmapfile{+sansmathaccent.map}
\def \ourFigPath {../../} 
\def \ourTablePath {../../Tables/}

\setbeamersize{text margin left=5mm,text margin right=12mm} 
%
\font\reali=msbm10 at 12pt
% subsets of real numbers
\newcommand{\numberset}{\mathbb}
\newcommand{\real}{\hbox{\reali R}}
\newcommand{\N}{\numberset{N}}
\newcommand{\realp}{\hbox{\reali R}_{\scriptscriptstyle +}}
\newcommand{\realpp}{\hbox{\reali R}_{\scriptscriptstyle ++}}
\newcommand{\virgolette}[1]{``#1''}
%

\author[Brianti, Gati]{Marco Brianti and Laura Gati}

\institute[Boston College]{Boston College}

\title{IT Spillovers in Long-Run TFP \\ A SVAR Approach}

\date{April 26, 2018}

\usetheme{Warsaw}

\begin{document}

\begin{frame}

\maketitle

\end{frame}

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Motivation}

	\begin{itemize}

	\item The Great Recession 2008 revived interest in endogenous components in TFP for explaining long-run TFP fluctuations.

	\

	\item[] $\rightarrow$ We explore the possibility that general-purpose technologies (GPT), in particular for the last 30 years information technologies (IT), play a role in explaining TFP fluctuations at long horizons.

	\

	\item However, an increase in IT investment affects TFP with some delays. Therefore this shock is likely to look very similar to a standard news shock.

	\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Motivation}

	\begin{itemize}
	\item Our main analysis is estimating a structural VAR in which we identify a shock to the exogenous component of future TFP (hereafter \textit{news shock}) and a contemporaneous shock which incentivizes investment in information and technology stock (hereafter \textit{IT shock}).

	\

	\

    \item As a support for our identification we design a 2-sector general equilibrium model which features both investment in IT and news shocks.

	\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
\frametitle{Preview of the Results}

\begin{itemize}
	\item IT shocks drive about half of long-run TFP fluctuations in our sample.

	\

	\item News shocks drive approximately one fifth $\Rightarrow$ Still relevant but - once appropriately cleaned from IT shocks - less important then found previously in the literature.

	\

	\item Our structural model points out that relative prices can be used to disentangle the two.

\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{VARs - A quick recap}
	\label{identification}

Structural Form	
\begin{equation}
(\mathbf{AD})^{-1}
    \mathbf{X_{t}}
= \mathbf{C(L)} 
\mathbf{X_{t-1}}
+ \mathbf{s_t}
\end{equation}

\

\

Reduced Form
\begin{equation}
\mathbf{X_{t}}
= \underbrace{\mathbf{AD} \mathbf{C(L)}}_\text{$\mathbf{B(L)}$} 
\mathbf{X_{t-1}}
+ \underbrace{\mathbf{AD} \mathbf{s_t}}_\text{$\mathbf{i_t}$}
\end{equation}

\

\

\begin{itemize}
	\item $AD$ is the impact matrix
	\item $A$ is s.t. $As_t s_t' A' = i_t i_t' = \Sigma$ and $s_t s_t' = I$
	\item $D$ is a rotation matrix s.t. $DD' = I$ $\Rightarrow$ $AD(AD)' = \Sigma$
\end{itemize}

We impose our identifying assumptions on the matrix $D$. \hyperlink{Technicalities}{\beamergotobutton{Technicalities}}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Related literatures}
	\label{related_lit}
\begin{itemize}
\item One strand of literature: Exogenous TFP and news shocks
	\begin{itemize}
	\item Beaudry \& Portier (2006)
	\item Barsky \& Sims (2011)

	\end{itemize}
\item [] \textbf{Our contribution: allow in this setting the existence of an endogenous mechanism that affects future TFP}

	\

	\

\item Another strand of literature: Endogenous TFP with R\&D investment as the key variable
	\begin{itemize}
	\item Comin \& Gertler (2006)
	\item Moran \& Queralto (2017)
	\item Guerron \& Jinnai (2014)

	\end{itemize}

\item []  \textbf{Our contribution: provide what we think is a more convincing test for the endogenous mechanism}

\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Identification: relative prices}

Let $P$ be the price level (price of final good), $P^{IT}$ the price of the IT good and relative prices be $P^{IT}/P$.

\

Intuition:
\begin{itemize}
\item As long as the technology improvement heralded by the news shock does not realize, relative prices should not move.
\item An increase in IT productivity shock makes only the IT sector more productive $\rightarrow$ impacts the sectoral price $P^{IT}$ \textbf{only} $\Rightarrow$ relative prices move! 
\end{itemize}

\

$\Rightarrow$ identifying restriction: news shock does not move relative prices.
\newline Accounting for price rigidities: news shock does not move relative prices after price adjustment has taken place (6-12 quarters).

\

Similar identification scheme: Fisher 2006.

%The price level $P$ is an aggregate of the prices intermediate goods. These are determined as an FOC to 
%\begin{equation*}
%Y_t(s) = \Psi_t K_t(s)^{\alpha}L_t(s)^{1-\alpha}
%\end{equation*}
%
%\hspace{5cm} $\hookrightarrow P$ is a function of $\Psi_t$ 
%
%\
%
%The price of IT good, $P^{IT}$, is determined as an FOC to
%
%\begin{equation}
%V_{i,t} = \lambda_t \Psi_t f(S_{i,t})
%\end{equation}
%
%
%\hspace{5cm}  $\hookrightarrow P^{IT}$ is a function of $\Psi_t$ \textbf{and} $\lambda_t$	
%
%\
%
%$\Rightarrow$ The two shocks imply different responses of relative prices $\frac{P^{IT}}{P}$!
%
%\begin{itemize}
%\item A news shock moves both $P$ and $P^{IT}$, leaving relative prices unchanged.
%\item An IT productivity shock moves only $P^{IT}$, thus changing relative prices.
%\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Identification strategy overall}

Recall that the innovation in TFP at time $t$ is:

\begin{equation*}
\epsilon^{TFP}_t =   \underbrace{V_{t-k}}_\text{news shock} + \underbrace{f(IT_{t-j})}_\text{IT productivity shock}  + \underbrace{\varepsilon_t}_\text{surprise tech shock} 
\end{equation*}

\

    \begin{enumerate}
    	\item The news shock $V_{t-k}$ maximizes the FEV of future TFP, with a zero impact effect, subject to the restriction that it has no effect on the relative price $RP$ at a small number of quarters;

	\

    \item The IT productivity shock maximizes the remaining FEV of future TFP, with a zero impact effect;

    \

    \item The tech shock $\varepsilon_t$ is considered as a residual shock and is left unrestricted (unidentified).
   \end{enumerate}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Empirical analysis}

	We run a SVAR using aggregate, quarterly US data. The data vector is:

	\begin{equation}
	\mathbf{X_t} = 
	\begin{bmatrix}
    TFP_t      \\

   SP_t   \\

   IT_t \\

   GDP_t \\

   C_t \\

   RP_t
\end{bmatrix}
	\end{equation}

\

\

\begin{itemize}
\item $RP = \pi^{IT}/\pi^{CPI}$. 
\item All variables are real (except price indexes) and in log levels (except for RP, which is in growth rates). 
\item The dataset ranges from 1989:q1 - 2017:q2.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Our favorite specification}

	\begin{itemize}
	\item Recall: dataset is quarterly and covers 1989:q1-2017-q2.

	\

	\item One lag (as suggested by BIC and HQ).

	\

	\item Horizon of FEV-maximization: 60 quarters.

	\

	\item Restriction on relative prices after a news shock is imposed at 8 quarters.

	\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{TFP response to both shocks}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_News_Shock_on_TFP__} % REGENERATE TFP SHOCKS FIGURES WITH RIGHT SCALES!!
\end{figure}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_IT_Shock_on_TFP__}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Real SP500 response to both shocks}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_News_Shock_on_Real_SP__}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_IT_Shock_on_Real_SP__}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{IT investment response to both shocks}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_News_Shock_on_Real_IT_Investment__}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_IT_Shock_on_Real_IT_Investment__}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Other responses to both shocks}

\vspace{-0.2cm}
\begin{figure}
\begin{multicols}{2}
\centering 
\includegraphics[scale = 0.1]{\ourFigPath Figures/fig_News_Shock_on_Real_GDP_Ryan_two_stepsID_10-Dec-2017_17_34_47}\\ 
\vspace{0.3cm}
\includegraphics[scale = 0.1]{\ourFigPath Figures/fig_News_Shock_on_Real_Consumption_Ryan_two_stepsID_10-Dec-2017_17_34_50}\\ 
\vspace{0.3cm}
\includegraphics[scale = 0.1]{\ourFigPath Figures/fig_News_Shock_on_Relative_Price_Ryan_two_stepsID_10-Dec-2017_17_34_52}\\

\includegraphics[scale = 0.1]{\ourFigPath Figures/fig_IT_Shock_on_Real_GDP_Ryan_two_stepsID_10-Dec-2017_17_35_00}\\
\vspace{0.3cm}
\includegraphics[scale = 0.1]{\ourFigPath Figures/fig_IT_Shock_on_Real_Consumption_Ryan_two_stepsID_10-Dec-2017_17_35_02}\\ 
\vspace{0.3cm}
\includegraphics[scale = 0.1]{\ourFigPath Figures/fig_IT_Shock_on_Relative_Price_Ryan_two_stepsID_10-Dec-2017_17_35_04}\\

\end{multicols}
\end{figure}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{FEV explained by the two shocks at 60 periods}

    \hspace{2.25cm}
\begin{large}
	\begin{tabular}{lccc}
	\hline
		& News & IT & Total \\
		\hline
	TFP	           & 0.20384  & 0.52596 & 0.72981  \\
		\hline
	\end{tabular}
\end{large}

\vspace{2cm}

For BS, FEV of news was 45\%.

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Interpretation}

\begin{enumerate}
 \item Shape and timing of the responses reflect ...

 %\

	\begin{itemize}
	\item both the Barsky \& Sims result;

	\vspace{0.1cm}

	\item as well as the the conjecture that the IT shock looks similar to news along certain dimensions;

	\vspace{0.1cm}

	\item but relative prices do indeed introduce a margin of difference between the two shocks.
	\end{itemize}

\

\item Shares of FEV of TFP explained ...
	\begin{itemize}
	\item are also in line with the Barsky \& Sims result;
	\item [] (for BS, news explains around 45\%, compared to 20\% here)

	\vspace{0.1cm}

	\item  yet suggest that the IT shock plays an important role as well (around 52\%).

	\vspace{0.1cm}

	\item And indeed the IT shock \emph{complements} the news shock, instead of substituting for it.
	\item [] For BS, the single identified shock explains around 45\%, while our two identified shocks explain around 73\%.
	\end{itemize}

\end{enumerate}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Robustness checks}

\begin{itemize}
\item Different variables
	\begin{itemize}
	\item Add the Michigan index of consumer confidence (expected business conditions 5 years ahead)

	\

	\item Replace IT prices with capital prices (following Comin \& Gertler)

	\

	\item Replace CPI inflation with PCE inflation
	\end{itemize}

	\

\item Different horizons at which we impose the restriction on relative prices for the news shock
\item[] $\rightarrow$ ran  6, 8, 10, 12 and 16 quarters.

\

\item Increase the number of lags (2)

\

\item Check whether VAR is information-sufficient to identify the news shock (Forni-Gambetti test) (p-val of 12\%)
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{A counterfactual}

	\begin{figure}
		\centering
		\includegraphics[scale=0.28]{\ourFigPath Figures/fig_counterfactual_}
	\end{figure}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Conclusion}

\begin{itemize}
\item Provided a test for the role of IT as an example of GPTs in explaining fluctuations in long-run TFP, and in doing so, overcame an econometric challenge prevalent in the literature on long-run productivity.

\

\item The results show that by controlling for the presence of news shocks, IT productivity shocks are important drivers of TFP fluctuations at long horizons.

\

\item This result does not contrast however with the findings of the news shock literature since we still find that news also play a significant role in explaining TFP.

\

\item Moreover, IT productivity shocks can be thought of as giving more microfoundations to what classical news shocks carry information on.

\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Work ahead}

\begin{itemize}

\item Show the identification assumption in the structural model.

\

\item Technically, should do a VECM due to cointegration and wanting to have variables in the VAR as growth rates rather than levels...

\

\item Dig deeper: IT is just an example of GPTs for the last 30 years
\item[] $\hookrightarrow$ could redo analysis for different time periods with different GPTs (electricity in 1920s, airplane industry in 1960s ...)

\

\item Use a fully structural model to see if we can find interesting interactions between the two shocks?
\item [] $\hookrightarrow$ In particular, we're thinking of noise shocks on IT productivity

\

\item But there are also other shocks that are similar to news shocks and yet not news shocks in the Barsky \& Sims sense: reallocative shocks, shocks to inventories, etc...  
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%                     APPENDIX  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{The current TFP slowdown}

	\vspace{-1cm}
	\noindent
	\begin{figure}
		\centering
		\includegraphics[scale=0.28]{\ourFigPath Figures/fig_TFP_macrolunch_29-Nov-2017_17_41_17}
	\end{figure}

\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Barsky \& Sims FEV of TFP explained}
	\label{BS_FEV}

\vspace{-1cm}
\noindent
\begin{figure}
\centering
\includegraphics[scale=0.3]{\ourTablePath fig_BS_fevtable}
\end{figure}

\hyperlink{related_lit}{\beamerreturnbutton{Return}}	
\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Production of IT goods}
	\label{it_sector}

Production function for IT goods:

\begin{equation}
V_{i,t} = \lambda_t \Psi_t f(S_{i,t}) 
\end{equation}
where $S_{i,t}$ is investment by producer $i$ in IT goods and $\lambda$ is the productivity of the IT sector.

\

The IT producer's problem is to max discounted profits $J_{i,t+1}$

\begin{equation}
\max_{S_{i,t}} \; \mathbb{E}(\Lambda_{t,t+1},J_{i,t+1})\lambda_t \Psi_t f(S_{i,t}) - P^{IT}_t S_{i,t}
\end{equation}

\
FOC:

\begin{equation}
P^{IT}_t = \mathbb{E}(\Lambda_{t,t+1},J_{i,t+1})\lambda \Psi_t f_1
\end{equation}

\hyperlink{expanding_variety}{\beamerreturnbutton{Return}}	
\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Timing: RD drop vs TFP drop}
	\label{timing}

	\vspace{-1cm}
	\noindent
	\begin{figure}
		\centering
		\includegraphics[scale=0.28]{\ourFigPath Figures/fig_RD_level_macrolunch_30-Nov-2017_11_31_04}
	\end{figure}

\hyperlink{convincing}{\beamerreturnbutton{Return}}	
\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{IT investment: a drop at the right time}
	\label{it_investment}

%	\vspace{-1cm}
	\noindent
	\begin{figure}
		\centering
		\includegraphics[scale=0.29]{\ourFigPath Figures/fig_IT_level_macrolunch_30-Nov-2017_11_37_26}
	\end{figure}

\hyperlink{convincing}{\beamerreturnbutton{Return}}	
\end{frame}
%%%%%%%%%%%%%%%%%

%%%%%%% Slide %%%%%%
\begin{frame}
\frametitle{Identification Strategy}
\label{Technicalities}

\begin{equation}
D = \begin{bmatrix}
d_{11} & \gamma_{12} & \gamma_{13} & d_{14} & \cdots \\
d_{21} & \gamma_{22} & \gamma_{23} & d_{24} & \cdots \\
\vdots & \vdots & \vdots & \ddots & \vdots 
\end{bmatrix}
\end{equation}

\begin{itemize}
	\item Indifferent over $d_{ij}$ as long as $D$ is orthogonal
	\item $A \gamma_{2}$ is the impact response to a news shock
	\item $A \gamma_{3}$ is the impact response to a IT productivity shock
	\item First element of both $A \gamma_{2}$ and $A \gamma_{3}$ is zero due to the no-contemporaneous effect of both shocks on TFP
	\item $A \gamma_2$ is such that the FEV of TFP is maximized subject to zero long-run effect on $RP$
	\item $A \gamma_3$ is maximizing the remaining FEV of TFP 
\end{itemize}

\hyperlink{identification}{\beamerreturnbutton{Return}}	
\end{frame}
%%%%%%%%%%%%%%%%%

\end{document}