### 2019-03-06 Latex/Conference_Modena_March2019/Brianti Gati March 2019.tex
slides for Modena presentation

\documentclass[hyperref={pdfpagelabels=false}]{beamer}
\setbeamercolor{background canvas}{bg=white}
\usepackage{graphicx,lmodern,subfigure,ulem,color,graphicx,tikz,booktabs,natbib}
\usepackage{mathrsfs}
\usetheme{Warsaw}
%\definecolor{beamer@blendedblue}{rgb}{0.1,0.5,0.1}
%\definecolor{ForestGreen}{RGB}{60, 140, 60}
%\setbeamercolor{structure}{fg=beamer@blendedblue}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]
\bibliographystyle{chicago}
\newcommand{\spitem}{\vspace{.3cm}\item}
\newcommand{\elas}{$E_{labor}$}
%\def \FigPath {Users\th3\Documents\Job_Market_Paper\Code\Figures} 
\DeclareMathOperator{\E}{\mathbb{E}}
\usepackage{makecell}

\title{ICT-Specific Investment Shocks and Economic Fluctuations}
%\\ \bigskip \\ \normalsize{Evidence and Theory of a General-Purpose Technology}
\author{Marco Brianti, Boston College\\Laura G\'ati, Boston College}
\institute{Convegno RECent\\Modena}
\date{March 2019}

\usetheme[
outer/progressbar=foot,
outer/numbering=none
]{metropolis}

\begin{document}

	\frame{\titlepage}

		\frame{\frametitle{Motivation}

	Why did U.S. output grow so slowly since the recession trough in 2009?

	\

	Fernald, Hall, Stock and Watson (2017) empirically show that two components explain nearly all this growth gap
	\begin{enumerate}
		\item Slow growth in total factor productivity (TFP)
		\item Falling labor force participation
	\end{enumerate}

\

In this paper we focus on the slowdown in TFP,
\begin{itemize}
	\item[$\Rightarrow$] Which variable(s) actually boost productivity? Why? 
\end{itemize}

	}

	\frame{\frametitle{An alternative to R\&D Investment}

		Following Romer (1990) and Coming and Gertler (2006), several papers endogenize TFP through \textbf{R\&D investment}
		\begin{itemize}
			\item Financial crisis $\Rightarrow$ $\downarrow$ $Y^D$ $\Rightarrow$ $\downarrow$ R\&D $\Rightarrow$ $\downarrow$ TFP $\Rightarrow$ $\downarrow$ $Y^S$
		\end{itemize}

		\

		Unfortunately, as showed by Oliner et al. (2007) and Jorgerson et al. (2008), TFP have started slowing down \textbf{before} the crisis.
		%\begin{itemize}
		%	\item[$\Rightarrow$] Medium-run business cycle models cannot be the full story
		%\end{itemize}

	\

	We follow an alternative avenue and focus on the relation between \textbf{information and communication technology investment} (ICTI) and future total factor productivity
	\begin{itemize}
			\item \textit{Expenditure in equipment and computer software meant to be used in production (as an input) for more than a year.}
			\end{itemize}

}

\frame{\frametitle{Our Contribution}

	\begin{enumerate}
		\item \textbf{Empirical evidence} that contemporaneous jumps in ICTI explain significant and persistent increases in future TFP
		\begin{itemize}
			\item SVAR techniques on U.S. aggregates to identify supply shocks specific to the ICT sector
		\end{itemize}

	\

	\

		\item We analyze a \textbf{GE model} to rationalize our empirical findings and draw conclusions concerning the nature of ICT 
		\begin{itemize}
		\item We extend the two-sector model of Greenwood et al. (1997) to accommodate for the role of ICT capital in production
		\end{itemize}
	\end{enumerate}

}

	\frame{\frametitle{Information and Communication Technology}

		ICT Investment is defined as \textit{total expenditure in equipment and computer software meant to be used in production (as an input) for more than a year.}
		\begin{itemize}
			\item $\uparrow$ ICT Investment  $\Rightarrow$ $\uparrow$ ICT capital stock in the economy
		\end{itemize}

\

\

	}

	\frame{\frametitle{Motivation: Empirical Challenge in Structural VAR}

	Empirically distinguishing between financial and uncertainty shocks is difficult
	\begin{itemize}
		\item[$\Rightarrow$] financial distress is empirically associated with larger volatility 
	\end{itemize}

\

Within a SVAR framework, this correlation significantly complicates identification of both shocks
\begin{enumerate}
	\item Implausible \textbf{zero-contemporaneous restrictions}
	\begin{itemize}
		\item[$\Rightarrow$] Both $F_t$ and $U_t$ are fast moving
	\end{itemize}	
	\item Unavailable instruments for \textbf{sign restrictions}	
	\begin{itemize}
		\item[$\Rightarrow$] Current theoretical models predict same qualitative effects on both prices and quantities
	\end{itemize}
\end{enumerate}

}

\frame{\frametitle{My contribution}

	I want to take a step back and show evidence and theory that financial and uncertainty shocks are \textbf{qualitative different}.

	\

	In particular,

	\begin{enumerate}
		\item \textbf{Corporate cash holdings} respond differently to financial and uncertainty shocks. 
		\begin{itemize}
			\item[$\Rightarrow$] Identification assumption
			\end{itemize}

		\

		\item I provide a \textbf{new econometric tool} to simultaneously identify two structural shocks when an internal instrument is available.\begin{itemize}
			\item[$\Rightarrow$] Generalized Penalty Function Approach  
		\end{itemize}

	\end{enumerate}
}

\frame{\frametitle{Roadmap}
	\Large{
		$ \ \ \ \ \ $ 1. \textbf{Cash Holdings}

		$ \ \ \ \ \ $ 2. Model

		$ \ \ \ \ \ $ 3. Empirical Strategy

		$ \ \ \ \ \ $ 4. Results

		$ \ \ \ \ \ $ 5. Conclusions
	}
}

\frame{\frametitle{Corporate Cash Holdings}

	\

\textbf{Cash and Cash Equivalents} refer to assets a business holds as ready cash
\begin{itemize}
	\item Coffer as petty cash
	\item Bank accounts
	\item Certificates of deposits
\end{itemize}

\

U.S. large firms have cash equal to about 15\% of total assets.

\

It is a \textbf{stock variable}, 
$$
Cash_t = Cash_{t-1} + NY_t + \delta K_t - I_t + B_t - D_t.
$$

}

\frame{\frametitle{Cross-Sectional Evidence}

	\textbf{Cash} and \textbf{Financial Frictions}
	\begin{itemize}
		\item[$\Rightarrow$] Cash is a substitute for external finance 
		\item[] \textit{Kaplan and Zingales (1997); Almeida, et al. (2004); Campello at al. (2010); Campello et al. (2011).}
	\end{itemize}

\

\textbf{Cash} and \textbf{Uncertainty}
	\begin{itemize}
	\item[$\Rightarrow$] Cash is positively associated with uncertainty shocks 
	\item[] \textit{Han and Qiu (2007); Baum et al. (2008); Bloom et al. (2018); Alfaro et al. (2018).}
\end{itemize}

}

\frame{\frametitle{Aggregate Evidence}

Aggregate quarterly cash (CHEQ) and assets (ATQ) using \textbf{Compustat} from 1961 to 2018.

\

Remove seasonality using 7-term Henderson filter on aggregate cash and aggregate assets and obtain \textbf{Cash2Assets}.

\

	\begin{table}[h!]
	\centering
	\begin{tabular}{llcc} \hline \hline
		                         & $\Delta$GDP         & U                &                   F    \\ 
		\hline
		\textit{Correlations}    &                     &                  &                                            \\ 
		U                        &      $-0.48^{***}$                &                  &                    \\
		F                        &  $-0.36^{***}$   &     $0.22^{***}$       &                             \\ 
		C2A                      &     $-0.06  $               &       $0.43^{***}$      &            $-0.37^{***}$           \\ 
		\hline \hline
	\end{tabular}
\end{table}

}

\frame{\frametitle{Roadmap}
	\Large{
		$ \ \ \ \ \ $ 1. Cash Holdings

		$ \ \ \ \ \ $ 2. \textbf{Model}

		$ \ \ \ \ \ $ 3. Empirical Strategy

		$ \ \ \ \ \ $ 4. Results

		$ \ \ \ \ \ $ 5. Conclusions
	}
}

%\frame{\frametitle{Economic Intuition I}

%To provide an economic intuition of the differential response of \textbf{cash holdings} to uncertainty and financial shocks, I present a properly augmented model in the spirit of 
%\begin{itemize}
	%\item Almeida, Campello, and Weisbach (2004)
	%\item Han and Qiu (2007)
%\end{itemize}

%\

%It is a simple representation of a dynamic setting where a credit-constrained profit-maximizing firm has a trade-off between present and future investment opportunities

%}

\frame{\frametitle{Model - General Setup}

\begin{itemize}

\item \textbf{Three-period partial equilibrium} model

\

\item Firm \textbf{maximizes} sum of \textbf{dividends} 
\begin{itemize}
\item Discount factor $\beta$ is one
\end{itemize}

\

\item \textbf{Choice variables} are
\begin{itemize}
	\item \textbf{Investments} $i_0$ and $i_1$ in period 0 and 1
	\item Amount to \textbf{borrow} $b_0$ and $b_1$ in period 0 and 1
	\item \textbf{Cash} $c$ in period 0 to be carried in period 1
\end{itemize}

\

\item Feature \textbf{financial frictions} in the form of risk premium

\

\item \textbf{Gross returns} $g(\cdot)$ happen in the last period for both investments
\begin{itemize}
	\item where $g'(\cdot) > 0$ and $g''(\cdot) < 0$.
\end{itemize}

	\end{itemize}

}

\frame{\frametitle{Model - Analytical Setup}

\begin{itemize}
	\item[]
\begin{itemize}
     \item[Period 0] $ \ \ \ d_0 = y_0 + b_0 - i_0 - c$

	\

	\item[Period 1] $ \ \ \ d_1 = y_1 + b_1 - i_1 + c, \ \ \ $ where $ \ y_1 \sim F(y_0,\sigma^2)$

	\

	\item[Period 2] $ \ \ \ d_2 = g(i_0) - b_0(1 + r_0) + g(i_1) - b_1(1 + r_1)$
\end{itemize}
\end{itemize}

\begin{eqnarray*}
	\begin{aligned}
\max_{\{ b_t,i_t,c \}_{t=0,1}} \ \ &\E \Big[  d_0 + d_1 + d_2     \Big| F  \Big] \\
\text{subject to} \ \ &r_0 = \frac{1}{2}\alpha_0 b_0 \ \ \text{and} \ \ r_1 = \frac{1}{2}\alpha_1 b_1 \\
&d_t \geq 0, \ \ \ t = 0,1,2 \\
\end{aligned}
\end{eqnarray*}

\

\begin{center}
Financial shock: $\uparrow \alpha_0$ $ \ \ \ \ $ vs $ \ \ \ \ $ Uncertainty shock: $\uparrow \sigma^2$
\end{center}

}

\frame{\frametitle{Solution}

	Assuming that firm needs external finance in equilibrium, model implies: 
\begin{itemize}
	\item $i_0 = y_0 + b_0 - c$,
	\item $i_1 = y_1 + b_1 + c$,
\end{itemize}
and first order conditions are
\begin{itemize}
	\item[$b_0:$] $g'(y_0 + b_0^* - c^*) = \underbrace{1 + \alpha_0 b_0^*}_{\text{Marginal Cost of $i_0$}}$

	\

	\item[$b_1:$] $\E \Big[ g'(y_1 + b_1^* + c^*) \Big] = \underbrace{1 + \alpha_1 b_1^*}_{\text{Marginal Cost of $i_1$}}$

	\

	\item[$c:$] $\underbrace{\E \Big[ g'(y_1 + b_1^* + c^*) \Big]}_{\text{Expected Marginal Return of $i_1$}} = \underbrace{g'(y_0 + b_0^* - c^*)}_{\text{Marginal Return of $i_0$}}$
\end{itemize}

}

%$\E_0 \Big[ g(y_t) \Big] > 1$ for $t=0,1$
%\begin{itemize}
	%\item[$\Rightarrow$] $d_t = 0 \ \ $ for $ \ t=0,1$
%\end{itemize}

%Objective function is,
%\begin{eqnarray*}
	%\begin{aligned}
	%	\max_{b_0,b_1,c} \ \ &g \big( i_0 \big) - b_0 - \frac{1}{2} \alpha_0 b_0^2 - y_0 + \E \bigg[  g \big( i_1 \big) - b_1 - \frac{1}{2} \alpha_1 b_1^2 - y_1   \bigg| F     \bigg]
	%\end{aligned}
%\end{eqnarray*}

\frame{\frametitle{Comparative Statics}

\begin{itemize}
	\item[$b_0:$] $g'(y_0 + b_0^* - c^*) = \underbrace{1 + \alpha_0 b_0^*}_{\text{Marginal Cost of $i_0$}}$

	\

	\item[$b_1:$] $\E \Big[ g'(y_1 + b_1^* + c^*) \Big] = \underbrace{1 + \alpha_1 b_1^*}_{\text{Marginal Cost of $i_1$}}$

	\

	\item[$c:$] $\underbrace{\E \Big[ g'(y_1 + b_1^* + c^*) \Big]}_{\text{Expected Marginal Return of $i_1$}} = \underbrace{g'(y_0 + b_0^* - c^*)}_{\text{Marginal Return of $i_0$}}$
\end{itemize}

\

\textbf{Uncertainty shock}: $y_1 \sim Q$ which is mean-preserving spread in $F$
	\begin{enumerate}
	\item[$\Rightarrow$] $c^*(\alpha_0,Q) > c^*(\alpha_0,F)$ as long as $g'''(\cdot) > 0$
\end{enumerate}

\

\textbf{Financial shock}: $\alpha_0^f > \alpha_0$ which is an exogenous increase in $r_0$
	\begin{enumerate}
	\item[$\Rightarrow$] $c^*(\alpha_0^f,F) < c^*(\alpha_0,F)$ 
\end{enumerate}

}

\frame{\frametitle{Roadmap}
	\Large{
		$ \ \ \ \ \ $ 1. Cash Reserves

		$ \ \ \ \ \ $ 2. Model

		$ \ \ \ \ \ $ 3. \textbf{Empirical Strategy}

		$ \ \ \ \ \ $ 4. Results

		$ \ \ \ \ \ $ 5. Conclusions
	}
}

\frame{\frametitle{Empirical Analysis}

Given the reduced-form system $X_t = B(L) X_{t-1} + \iota_t$ where

$$
X_t = \begin{bmatrix}
U_t \\
F_t \\
GDP_t \\
C_t \\
I_t \\
H_t \\
C2A_t \\
GDPDef_t \\
\end{bmatrix}
$$

\

Dataset ranges from 1978q1 to 2015q3.

}

\frame{\frametitle{Objective of the Empirical Strategy}

Given the reduced-form system $X_t = B(L) X_{t-1} + \iota_t$,

$\Rightarrow$ find a rotation of $\Sigma_{\iota} = \iota_t' \iota_t$ such that

\begin{enumerate}
	\item it allows $F_t$ and $U_t$ to respond to both shocks on \textbf{impact}
	\item it respects \textbf{sign-restriction} assumptions on cash
	\item it is \textbf{unique}
	\item it delivers shocks \textbf{orthogonal} to each other
	\item it is unaffected by the \textbf{order} of the estimation
\end{enumerate}

}

\frame{\frametitle{Sequential Penalty Function Approach ($\delta \geq 0$)}

	\

	\textbf{1. Uncertainty Shock}
	\begin{eqnarray*}
		\begin{aligned}
		\max_{\gamma_{U}} \ \ \ \ \ &  \ \underbrace{e_U A_0 \gamma_U}_{\text{Impact on U}} \ \ \  + \ \ \ \textcolor{red}{\delta} \underbrace{e_{C} A_0 \gamma_U}_{\text{Impact on Cash}} \\
		\end{aligned}
	\end{eqnarray*}

\textbf{Intuition}. $\gamma_{U}$ increases both uncertainty and cash on impact.

\

	\textbf{2. Financial Shock}
	\begin{eqnarray*}
		\begin{aligned}
		\max_{\gamma_{F}} \ \ \ \ \ & \  \underbrace{e_F A_0 \gamma_F}_{\text{Impact on F}} \ \ \  - \ \ \  \textcolor{red}{\delta} \underbrace{e_{C} A_0 \gamma_F}_{\text{Impact on Cash}} \ \ s.t. \ \ \underbrace{\textcolor{red}{\gamma_U \gamma_F' = 0}}_{\text{Orthogonality with U shock}} \\
		\end{aligned}
	\end{eqnarray*}

\textbf{Intuition}. $\gamma_{F}$ increases uncertainty and decreases cash on impact.

}

\frame{\frametitle{Sequential Penalty Function Approach ($\delta \geq 0$)}

		\

	\textbf{1. Financial Shock}
	\begin{eqnarray*}
		\begin{aligned}
			\max_{\gamma_{F}} \ \ \ \ \ & \  \underbrace{e_F A_0 \gamma_F}_{\text{Impact on F}} \ \ \  - \ \ \  \textcolor{red}{\delta} \underbrace{e_{C} A_0 \gamma_F}_{\text{Impact on Cash}}  \\
		\end{aligned}
	\end{eqnarray*}

	\textbf{Intuition}. $\gamma_{F}$ increases uncertainty and decreases cash on impact.

	\

	\textbf{2. Uncertainty Shock}
	\begin{eqnarray*}
		\begin{aligned}
			\max_{\gamma_{U}} \ \ \ \ \ &  \ \underbrace{e_U A_0 \gamma_U}_{\text{Impact on U}} \ \ \  + \ \ \ \textcolor{red}{\delta} \underbrace{e_{C} A_0 \gamma_U}_{\text{Impact on Cash}} \ \ s.t. \ \ \underbrace{\textcolor{red}{\gamma_U \gamma_F' = 0}}_{\text{Orthogonality with F shock}} \\
		\end{aligned}
	\end{eqnarray*}

	\textbf{Intuition}. $\gamma_{U}$ increases both uncertainty and cash on impact.

}

\frame{\frametitle{Generalized Penalty Function Approach}

	\

	\textbf{1. Financial Shock}
	\begin{eqnarray*}
		\begin{aligned}
			\max_{\gamma_{F}} \ \ \ \ \ & \  \underbrace{e_F A_0 \gamma_F}_{\text{Impact on F}} \ \ \  - \ \ \  \textcolor{red}{\delta^*} \underbrace{e_{C} A_0 \gamma_F}_{\text{Impact on Cash}}  \\
		\end{aligned}
	\end{eqnarray*}

	\textbf{2. Uncertainty Shock}
	\begin{eqnarray*}
		\begin{aligned}
			\max_{\gamma_{U}} \ \ \ \ \ &  \ \underbrace{e_U A_0 \gamma_U}_{\text{Impact on U}} \ \ \  + \ \ \ \textcolor{red}{\delta^*} \underbrace{e_{C} A_0 \gamma_U}_{\text{Impact on Cash}}  \\
		\end{aligned}
	\end{eqnarray*}

where $\textcolor{red}{\delta^*}$ is chosen such that $\textcolor{red}{\gamma_U \gamma_F' = 0}$.

\

\textbf{Economic Intuition.} Weight of sign restrictions should be large enough such that the two shocks are separated without any external constraint.

}

%	\text{subject to} \ \ \ \ \  & \ \delta \geq 0 \ \ \ \ \ \ \text{and} \ \ \ \ \ \  \underbrace{\gamma_U \gamma_U' = 1}_{\text{Normalization}}

%		\text{subject to} \ \ \ \ \ \ & \ \delta \geq 0, \ \ \ \ \ \ \ \underbrace{\gamma_F \gamma_F' = 1}_{\text{Normalization}}, \ \ \ \ \ \ \text{and}

\frame{\frametitle{Roadmap}
	\Large{
		$ \ \ \ \ \ $ 1. Cash Reserves

		$ \ \ \ \ \ $ 2. Model

		$ \ \ \ \ \ $ 3. Empirical Strategy

		$ \ \ \ \ \ $ 4. \textbf{Results}

		$ \ \ \ \ \ $ 5. Conclusions
	}
}

\frame{\frametitle{Uncertainty Shock}

}

\frame{\frametitle{Financial Shock}

}

\frame{\frametitle{Variance Explained}

}

\frame{\frametitle{Roadmap}
	\Large{
		$ \ \ \ \ \ $ 1. Cash Reserves

		$ \ \ \ \ \ $ 2. Model

		$ \ \ \ \ \ $ 3. Empirical Strategy

		$ \ \ \ \ \ $ 4. Results

		$ \ \ \ \ \ $ 5. \textbf{Conclusions}
	}
}

\frame{\frametitle{Conclusions}

\begin{itemize}

\item \textbf{Cash reserves} as an internal instrument to simultaneously identify financial and uncertainty shocks.

\

\item An \textbf{econometric tool} to overcome known SVAR shortcomings
\begin{itemize}
	\item[$\Rightarrow$] Tests using simulated data confirm the reliability of the procedure. See \textbf{Appendix A}.
\end{itemize}

\

\item Empirical results confirm the \textbf{relevance} and \textbf{exogeneity} of both shocks.
\begin{itemize}
	\item[$\Rightarrow$] Correlations with external shocks is available in \textbf{Appendix B}.
\end{itemize}

\

\item \textbf{Financial shocks} have larger effects in the \textbf{short run} while \textbf{uncertainty shocks} have a more \textbf{persistent} effect.

\end{itemize}

}

\frame{\frametitle{Next Steps}

	\begin{enumerate}

\item Empirical evidence in favor of my identification assumption

\begin{itemize}
	\item Using \textbf{Quarterly Financial Report} data to show that my results are mostly driven by small firms 
	\item Merging \textbf{Compustat} and \textbf{TRACE} to show firm-level evidence of the differential response of cash
	$$
	\frac{Cash_{it}}{Assets_{it}} = \underbrace{\beta^U}_{(+)} U_{it} + \underbrace{\beta^F}_{(-)} F_{it} + \beta^X X_{it} + \delta_i + \lambda_t + \varepsilon_{it}
	$$
\end{itemize}

\

\item Design and analyze a \textbf{dynamic GE model} 
\begin{itemize}

	\item to show my identification assumption survives to GE effects

	\item to test whether GPFA can recover both shocks

\end{itemize}

\end{enumerate}

}

\frame{\frametitle{Appendix A - Simulated Data and Generalized PFA}

	Consider the following structural model,

	\begin{itemize}
		\item $U_t = B_{UU} U_{t-1} + B_{UF} F_{t-1}  + B_{UC} C_{t-1}  +  A_{UU}s_t^U  +  A_{UF}s_t^F$
		\item $F_t = B_{FU} U_{t-1} + B_{FF} F_{t-1}  + B_{FC} C_{t-1}  +  A_{FU}s_t^U  +  A_{FF}s_t^F$
		\item $C_t = B_{CU} U_{t-1} - B_{CF} F_{t-1}  + B_{CC} C_{t-1}  +  A_{CU}s_t^U  -  A_{CF}s_t^F$
		\end{itemize}
where $s_t^U \sim N(0, \sigma^2_U)$, $s_t^F \sim N(0, \sigma^2_F)$ and $s_t^U \perp s_t^F$.

\

	Objective is to estimate structural parameters 
	\begin{itemize}
		\item using only $X_t = [U_t, \ F_t, \ C_t]$, and
		\item only knowing that $A_{ji} \geq 0$ for $j,i = \{U, \ F, \ C \}$.
	\end{itemize} 
	$\Rightarrow$ apply GPFA to test reliability of the econometric tool
}

\frame{\frametitle{Appendix A - Small Sample Performance (T = 100)}

}

\frame{\frametitle{Appendix A - Large Sample Performance (T = 100000)}

}

\frame{\frametitle{Appendix B - Correlations with Other External Shocks}

	\begin{table}[h!]
		\centering
		\begin{tabular}{lrr} \hline \hline
			                & \textbf{Uncertainty Shocks} & \textbf{Financial Shocks} \\ 
			                \hline
			\textit{External Shocks} &                    &                      \\ 
			BZP Military News        & $-0.10$  $(0.24)$  &   $0.08$  $(0.31)$   \\ 
			Ramey Military news      &  $0.07$  $(0.44)$  &   $0.02$  $(0.82)$   \\ 
			LWY Exp. Tax             &  $0.03$  $(0.74)$  &   $0.15$  $(0.11)$   \\ 
			RRMR Unexp. Tax          & $-0.13$  $(0.16)$  &   $0.05$  $(0.59)$   \\ 
		    RRMR Exp. Tax            & $-0.08$  $(0.36)$  &   $0.03$  $(0.76)$   \\ 
		    AdjTFP AR(1)             &  $0.08$  $(0.31)$  &  $-0.14$  $(0.11)$   \\ 
		    RR Mon. Policy           & $-0.13$  $(0.18)$  &  $-0.04$  $(0.70)$   \\ 
		    \hline \hline
		\end{tabular}
	\end{table}

}

\end{document}